### 2016-04-15 guidelines.tex

\documentclass[10pt]{report}

\usepackage[T1]{fontenc}
\usepackage{amsmath}
\begin{document}
	\textbf{Information about the model code in MATLAB}\\

	1. Model generation is carried out from a excel/text file using modelgen.m\\

	2. Using FBA to calculate the steady state fluxes is optional (what if it is a toy model?)\\

		2a. If using FBA or PFBA, fix uptake fluxes, reaction bounds and essential and nonessential exchange reactions in the model.\\

		2b. For toy model steady state fluxes can be alternatively specified in the excel/text file\\

	3. Specify metabolite concentrations that you may know\\

		3a. Sample the other metabolite concentrations using ACHR within a constrained space.\\

		3b. If not sampling, assign metabolites a concentration either using a file or in the script file.\\

\end{document}