### 2018-06-07 ident/docs/draft_4.tex
new draft with corrections from WRC

\documentclass[10pt]{article}

\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{lmodern}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{pifont}
\usepackage{bm}
\usepackage{graphicx}
\usepackage[space]{grffile}
\usepackage{multicol}
\usepackage{array}
\usepackage{tabu}
\usepackage{ragged2e}
\usepackage{setspace}
\usepackage{xr} % package for linking external document references
\usepackage[font=small,labelfont=bf,labelsep=period]{caption}
%\usepackage{subcaption}
%\usepackage[CaptionAfterwards]{fltpage}
\usepackage{subfigure}
\usepackage{lineno}
\linenumbers
\usepackage{tikz}
\def\checkmark{\tikz\fill[scale=0.4](0,.35) -- (.25,0) -- (1,.7) -- (.25,.15) -- cycle;} 
\usepackage[margin=1.0in]{geometry}

\usepackage[backend=biber,style=authoryear,sorting=nyt,url=false,isbn=false,doi=false,firstinits=true]{biblatex}

\DeclareNameAlias{default}{last-first}

\DefineBibliographyStrings{english}{%
	andothers = {\addcomma\addspace\textsc{et\addabbrvspace al}\adddot},
	and = {\textsc{and}}
}
\renewcommand*{\labelnamepunct}{\space\space}

\renewbibmacro{in:}
{%
	\ifentrytype{article}{%
	}{%
		\printtext{\bibstring{in}\intitlepunct}%
	}%
}
\renewbibmacro*{volume+number}{%
	\printfield{volume}%
	\setunit*{\addcomma\space}%
	\printfield{number}%
	\setunit{\addcomma\space}}

\DeclareFieldFormat{pages}{#1}

\renewbibmacro*{publisher+location+date}{%
	\printlist{publisher}%
	\setunit*{\addcomma\space}%
	\setunit*{\addcomma\space}%
	\usebibmacro{date}%
	\newunit}

\renewcommand{\newunitpunct}{\addcomma\space}
\DeclareFieldFormat[article,inbook,incollection,inproceedings,patent,thesis,unpublished]{title}{#1} 
\DeclareFieldFormat{year}{#1}

\renewcommand{\baselinestretch}{2.0}
\addbibresource{refs/ident_refs.bib}
\captionsetup{font={stretch=2.0}}
\externaldocument{sifigures} % reference to existing external document

\begin{document}
	\begin{center}
		\begin{Large}
			Practical Identification and Experimental Design for Parameter Estimation in Kinetic Models of Metabolism
		\end{Large}\\
		Shyam Srinivasan\textsuperscript{a}, William R. Cluett\textsuperscript{a} and Radhakrishnan Mahadevan\textsuperscript{*,a,b}\\
	\end{center}
	a - Department of Chemical Engineering and Applied Chemistry, University of Toronto, Toronto, ON, Canada.\\
	b - Institute for Biomaterials and Biomedical Engineering, University of Toronto, Toronto, ON, Canada.\\
	{*} Corresponding author
	\section*{Abstract}
	\section{Introduction:}
	The use of metabolic engineering spans a wide variety of applications. Some notable examples include the design of microorganisms for the biosynthesis of commodity and specialty chemicals \parencite{Andreozzi2016}, engineering mammalian cells as therapeutic targets for cures to some ailments affecting humans \parencite{DiFilippo2016,Apaolaza2017}, and changing the constituents of the human gut microbial community to cure related diseases \parencite{Zerfab2018}. These applications require us to understand the numerous complex interactions, their roles in cell function, and sometimes even the mechanisms behind these interactions. Computational models offer a systematic way to integrate available experimental data, and to study and understand these interactions through mathematical representations of the biological systems in which these interactions occur \parencite{Bordbar2014a,Saa2017}. They are also used to predict changes in cell function based on changes in the type and nature of the modeled interactions \parencite{Andreozzi2016}, or aid in the identification of therapeutic targets for drug discovery and development \parencite{Bordbar2015,Chandrasekaran2017}

	Constraint-based models (CBMs) of metabolism are used to improve our understanding of metabolism by representing it as a stoichiometric network of reactions \parencite{Bordbar2014a}. The ability of CBMs to shine light on the nonintuitive interactions that govern cellular metabolism is leveraged to engineer and asses the impact of designs that alter the ability of a cell to grow, or produce a desired metabolite \parencite{Maia2016}. However, in CBMs, metabolism is assumed to operate under a pseudo steady state. Consequently, the metabolite concentrations within the metabolic network are assumed to be constant, and changes in metabolite concentrations are not modeled. Furthermore, since CBMs represent metabolism using only the stoichiometry of its constituent reactions, they do not account for the various non-catalytic regulatory interactions that are also responsible for metabolic function. These shortcomings prevent CBMs from being used to fully understand the steady state as well as the dynamic characteristics of metabolic networks.

	In contrast, the effects of regulatory interactions and changes in metabolite concentrations on different characteristics of metabolism can be studied using kinetic models of metabolism \parencite{Saa2017}. These models account for changes in metabolite concentrations subject to thermodynamic and regulatory constraints that underly metabolic networks in addition to their stoichiometry \parencite{Link2014}. Kinetic models can not only help us better understand lesser known and understood characteristics of metabolism like bistability \parencite{Kotte2014}, and their role in human health, but can also improve predictions about the impact of engineering design perturbations on metabolism, and help propose alternative designs to achieve metabolite production goals \parencite{Khodayari2016}.

	Kinetic models differ from CBMs in their use of mechanistic enzyme kinetics to model enzyme catalyzed fluxes within a metabolic network \parencite{Srinivasan2015,Saa2017}. The use of kinetic models requires information on the enzyme kinetic rate laws that will be used to model the fluxes within a metabolic network, as well as numerical values for the parameters used in these rate laws. Analyzing the ability of a metabolic network to exhibit dynamic characteristics like multiple steady states and oscillations, irrespective of the structure of the network, is one example where kinetic rate laws and parameter values play a crucial role \parencite{Srinivasan2017}.

	Despite their importance, the parameterization of kinetic models is still a problem for which solutions are a subject of debate within the modeling community. Typically, enzyme kinetic rate laws are parameterized based on in vitro observations of enzyme activity, as opposed to observations made under in vivo conditions \parencite{Heijnen2005,Smallbone2007}. However, some researchers have questioned their relevance for gleaning information on the dynamics of metabolism under in vivo conditions \parencite{Heijnen2005,Heijnen2013}. On the other hand, some reports have shown that despite the large uncertainties associated with parameters estimated based on in vivo experimental data \parencite{Link2014}, in vitro parameter estimates are a reasonable approximation of values that would be applicable under in vivo conditions \parencite{Davidi2016}.

	 Authors have sought to constrain and determine uncertainty in parameter estimates and associated model predictions by using Monte Carlo approaches for kinetic modeling of metabolism \parencite{Andreozzi2016a}. These approaches also allow for the integration of experimentally observed in vivo data. ORACLE \parencite{Wang2004} and Ensemble modeling \parencite{Tran2008}, are two such examples. These and other Monte Carlo kinetic modeling methods have been previously reviewed \parencite{Srinivasan2015}. 	 
	 Bayesian approaches to improve parameter estimation and quantify estimation uncertainty have also been proposed \parencite{Saa2016}. \textcite{Vanlier2013} provide a review of different Bayesian approaches to quantify parameter uncertainty.

	Related to the development of these methods to quantify parameter estimation uncertainty, the importance of model parameter identifiability, i.e., a necessary condition to estimate unique kinetic parameter values from experimental data, is often overlooked \parencite{Ljung1994,Berthoumieux2013}.

	Briefly, unique values can only be estimated for structurally identifiable parameters. If a parameter cannot be uniquely estimated due to its nonlinear relationships with other parameters (redundant model parameterization), or due to the inability to measure all possible metabolite concentrations (presence of unobservable system states), then the parameter is said to be structurally non-identifiable. Similarly, practical identifiability of a parameter concerns with the informativeness of the experimental data used for parameter estimation, and the ability to quantify any existing uncertainty in the parameter estimates. If the ability to estimate unique parameter values is compromised due to the inability of the available data to capture the requisite information needed to estimate the parameters in the modeled system, and the uncertainty in parameter estimates is not quantifiable, the parameter is said to be practically non-identifiable \parencite{Ljung1994,Rodriguez-Fernandez2006}.

	Authors have proposed ways to assess parameter identifiability by proposing approximate kinetic models of metabolism that utilize empirical enzyme kinetic rate laws with parameters that have physical significance, and are identifiable \parencite{Heijnen2005,Smallbone2007}. Significant work has also been done towards the development of methods for structural identification of parameters in kinetic models of metabolism \parencite{Ljung1994,Nikerel2009,Berthoumieux2013,Raue2014}. \textcite{Chis2011b} provide a comparison of the computational performance and implementation aspects of some of these methods.

	Methods to improve practical identifiability through a priori experimental design have also been developed, with a focus on kinetic models of metabolism \parencite{Gadkar2005a,Rodriguez-Fernandez2006,Vanlier2014a,Raue2014}. Some of these methods are limited by their applicability to approximate kinetic models only \parencite{Nikerel2009,Berthoumieux2013}, while some others suffer from computational limitations when applied to kinetic models of large metabolic networks \parencite{Gadkar2005a,Raue2014}.

	Most of the aforementioned methods assess identifiability (structural or practical) in the presence of dynamic time course concentration data. In this paper, we show that the identifiability of parameters for each individual reaction in kinetic models of metabolism can be done using just steady state fluxomics, metabolomics and proteomics data, and propose a methodology to do so. We also demonstrate how this method can also facilitate the design of experiments for generating the necessary steady state data to estimate unique parameter values for all reaction fluxes in a metabolic network. In doing so, we are able to propose the number and types of perturbations that will provide the most useful data for parameter estimation. We also assess the identifiability of different enzyme kinetic rate laws that are typically used to model fluxes in metabolic networks using steady state data. We illustrate our methodology using a small metabolic network model of glucoeneogenesis in \textit{Escherichia coli} \parencite{Kotte2014, Srinivasan2017} under the assumption that all intracellular metabolite concentrations and fluxes can be measured.

	\section{Methods}\label{sec:methods}
	In Section \ref{sec:kinetic_model} we present the typical structure of a kinetic model of metabolism and a preliminary description of the least squares method by which the model parameters are estimated from experimental data. In Section \ref{sec:ident_def} we provide formal mathematical definitions for structural and practical identifiability. The computer-algebra system based method to assess identifiability that we have developed is described in Section \ref{sec:ident} that follows. We also provide a description of a numerical equivalent for the computer algebra-based method to estimate parameters numerically in Section \ref{sec:numerical_method}. In Section \ref{sec:degree_of_identifiability}, we define a quantitative metric to describe the identifiability of parameters, and in Section \ref{sec:small-model} a complete description of the small metabolic network used to demonstrate the methodology that we have developed is provided. Finally, a description of how the method can be used for experimental design for parameter estimation is given in Section \ref{sec:experimental_design}.

	\subsection{Parameter estimation for kinetic models of metabolism}\label{sec:kinetic_model}
	In kinetic models of metabolism, ordinary differential equations (ODE) are used to express the rate of change of metabolite concentrations ($x\in\mathbb{R}^{n_x}$) as a function of the reaction fluxes ($v\in\mathbb{R}^{n_r}$) in the metabolic network (Equation \ref{eq:kinstoich}). The matrix $\mathbf{S}\in\mathbb{R}^{n_x \text{x} n_r}$ in Equation (\ref{eq:kinstoich}a) defines the stoichiometric relationship between the fluxes and the concentrations of the metabolic network.
	\begin{subequations}\label{eq:kinstoich}
		\begin{align}
		\dot{x} = \mathbf{S}v\\
		v = f(x, \theta, u)
		\end{align}
	\end{subequations}
	The expression for the nonlinear function ($f$) used to describe each reaction flux $v_i$ in $v$, $i={1, 2, ..., n_r}$, in a particular kinetic model (Equation \ref{eq:kinstoich}b) is dependent on the enzyme kinetic mechanism that is used to model the reaction \parencite{Srinivasan2015}. Accordingly, $f$ is typically a nonlinear function of the vector of metabolite concentrations ($x\in \mathbb{R}^{n_x}$), the vector of enzyme kinetic parameters ($\theta\in\mathbb{R}^{n_p}$) and other input concentrations ($u \in \mathbb{R}^{n_u}$).

	Parameter estimation methods based on optimization principles are used to determine parameter estimates from experimental data. Under the assumption that all intracellular metabolite concentrations and fluxes can be measured, a parameter estimation problem can be formulated as a nonlinear programming problem (Equation \ref{eq:lsqopt}) to estimate the values of enzyme kinetic parameters, $\theta \in \mathbb{R}^{n_p}$, based on the measured data. 
	\begin{subequations}\label{eq:lsqopt}
		\begin{align}
		\underset{\theta}{\mathrm{min}} &\sum_{k=1}^{n_x + n_r}\sum_{l=1}^{n_t}\left(\frac{y_{kl}^*-y_{kl}}{\sigma_{kl}^*}\right)^2\\
		&\theta_l \le \theta \le \theta_u
		\end{align}
	\end{subequations}
	Here $y \in \mathbb{R}^{n_x + n_r}$ is the combined vector of concentrations ($x\in \mathbb{R}^{n_x}$) and fluxes ($v \in \mathbb{R}^{n_r}$), at each time point $l={1, 2, ..., n_t}$. The minimization of the least squares error between the measured ($y^*$) and modeled ($y$) concentrations and fluxes, weighted by the variance in the experimental data $\sigma_{kl}^*$ for each concentration and flux, at each time point $l={1, 2, ..., n_t}$, is used as an objective function (Equation \ref{eq:lsqopt}a) for the optimization problem. The least squares parameter values are determined within fixed upper ($\theta_u$) and lower ($\theta_l$) bounds (Equation \ref{eq:lsqopt}b).

	Typically, not all metabolite concentrations and fluxes can be measured. Under these conditions, the formulation of the kinetic model and the parameter estimation problem shown in Equations (\ref{eq:kinstoich}) and (\ref{eq:lsqopt}) is different. These formulations are ubiquitously presented in other papers \parencite{Gadkar2005a,Rodriguez-Fernandez2006}. Hence, these will not be discussed in this paper.

	\subsection{Structural and practical identifiability of parameters in kinetic models}\label{sec:ident_def}
	In the Introduction, we briefly mentioned that the ability to estimate unique parameter values from available experimental data is governed by the identifiability of these parameters in the model \parencite{Ljung1994,Vanlier2012,Berthoumieux2013,Raue2014}. Below, we provide a formal definition of structural and practical identifiability of parameters.

	The parameters in $\theta$ in any model (Equation \ref{eq:kinstoich}) are said to be structurally identifiable if, for an input-output mapping defined by $y = \Phi(\theta,u)$ for at least one input function $u$, any two values of parameters $\theta_1$ and $\theta_2$ satisfy the relationship in Equation (\ref{eq:stident}):
	\begin{align}\label{eq:stident}
	\Phi(\theta_1,u) = \Phi(\theta_2,u) \iff \theta_1 = \theta_2
	\end{align}
	Accordingly, if parameters in $\theta$ have a unique value, a finite number of non-unique values or an infinite number of values for all input functions, they are said to be structurally globally identifiable, locally identifiable or non-identifiable, respectively. So first, structural identifiability of parameters in a dynamic model helps establish the presence or absence correlations between different model parameters \parencite{Rodriguez-Fernandez2006}. Second, in instances where not all concentrations and fluxes can be measured, structural identification can also elucidate relationships between the unmeasurable and measured concentrations/fluxes. Consequently, the effect of model structure and parameterization on the ability to infer true parameter values from experimental data is determined by the structural identifiability of the parameter.

	Experimental data from many physical systems is usually noisy, and sometimes may not contain all requisite information pertaining to the system. When parameters are estimated on the basis of this data, the ability to estimate unique parameter values to satisfy Equation (\ref{eq:stident}) is referred to as practical identifiability. If a unique parameter value satisfying Equation (\ref{eq:stident}) can be estimated, then $\theta$ is said to be globally practically identifiable. Whereas, if parameter estimates with quantifiable uncertainties can be found, then $\theta$ is said to be locally identifiable. No unique parameter estimates can be found if $\theta$ is practical non-identifiable. The practical identifiability of a parameter is hence contingent upon the nature, quality and quantity of data available to estimate the parameter as opposed to the structure and parameterization of the model.

	Therefore, in conclusion, establishing the structural identifiability of parameters enables one to propose models that are not only appropriate representations of physical processes, but are also parameterized in such a way that the value of these parameters can be estimated from measurable data. On the other hand, establishing practical identifiability of parameters in any model helps design experiments that are minimal, informative and useful for parameter estimation.

	\subsection{A method to determine structural and practical identifiability of kinetic models of metabolism}\label{sec:ident}
	In this section, we provide the mathematical framework for \textcolor{red}{identification} of parameters in kinetic models of metabolism. A summary of the methodology in the form of a flow diagram is shown in Figure \ref{fig:ident-flowchart}. As indicated in Figure \ref{fig:ident-flowchart}a, the first step involves the construction of the kinetic model (Equation \ref{eq:kinstoich}) of the metabolic network with $n_r$ reaction fluxes.

	\begin{figure}[!tbhp]
		\centering{\includegraphics[width=.9\textwidth,height=.6\textheight,keepaspectratio]{figures/figure3/ident_analysis}}
		\caption{A flow diagram showing the methodology developed to establish practical identifiability of parameters in kinetic models of metabolism. a) The steps for the construction of a kinetic model of a metabolic network. The choice of rate law formulations to describe metabolic fluxes influences the identification methodology. The identifiability of parameters for each flux can be established independently. b) The steps for practical identifiability analysis for parameters of a single flux.}\label{fig:ident-flowchart}
	\end{figure}

	For each flux $v_i$, $i={1, 2, ..., n_r}$, in the kinetic model, let $\mathbf{\theta} \in \mathbb{R}^p$ in Equation (\ref{eq:kinstoich}b). If data from $n_E$ experiments is available for the chosen metabolic network, as stated earlier, for each experiment $j = {1, 2, ..., n_E}$, we assume that all metabolite concentrations ($x\in\mathbb{R}^{n_x}$) and reaction fluxes ($v\in\mathbb{R}^{n_r}$) are measurable. We discuss the implications of relaxing this assumption in the results section. The pertinent information for each experiment $j$ is available as a vector of concentrations and fluxes, $\mathbf{x}_j$ and $\mathbf{v}_j$, respectively (Figure \ref{fig:ident-flowchart}b).

	In order to establish the practical identifiability of kinetic parameters for each flux $v_i$, $i={1, 2, ..., n_r}$, we describe a computer algebra-based method. The primary use of the computer algebra system is to obtain closed-form expressions for each parameter in $\mathbf{\theta} \in \mathbb{R}^p$ for each flux $v_i$ (Figure \ref{fig:ident-flowchart}b). This is done by first selecting a combination of $p\le n_E$ experimental data. The fluxes and concentrations from $p$ different experiments are then used to formulate a system of nonlinear algebraic equations in $\mathbb{R}^p$ for each flux $v_i$, shown in Equation (\ref{eq:nonlineq}). 
	\begin{align}\label{eq:nonlineq}
	v_{i, j} = f_j(\mathbf{x}_j,\mathbf{\theta}, \mathbf{u}_j) && \forall j=\{1, 2, ..., p\}\subset\{1, 2, ..., n_E\}
	\end{align}
	Here, $v_{i,j}$ refers to the flux $v_i$ obtained from experiment $j$. $\mathbf{x}_j$ and $\mathbf{u}_j$ are the vector of metabolite and other input concentrations from each experiment $j$, and $\mathbf{\theta}$ is a vector in $\mathbb{R}^p$, whose elements are denoted by $\theta_k$.

	Each equation in (\ref{eq:nonlineq}), indicated by the index $j$, corresponds to the kinetic rate law expression $f(x, \theta, u)$ for each $v_i$, $i={1, 2, ..., n_r}$, described in Equation (\ref{eq:kinstoich}b), written for concentrations ($\mathbf{x}_j$, $\mathbf{u}_j$) and fluxes ($v_{i,j}$) obtained from experiment $j$. Solving the system in Equation (\ref{eq:nonlineq}) results in $\mathbb{R}^p$ nonlinear expressions for each parameter $\theta_k$ in $\theta \in \mathbb{R}^p$ (Equation \ref{eq:theta-eq}), where $N(\mathbf{v}_i, \mathbf{x}, \mathbf{u})$ is the numerator of $g$, and $D(\mathbf{v}_i, \mathbf{x}, \mathbf{u})$ is the denominator of $g$ (Figure \ref{fig:ident-flowchart}b). Note that $\mathbf{v}_i$, $\mathbf{x}$ and $\mathbf{u}$ are used to denote vector of vectors of fluxes for reaction $i$ ($\mathbf{v}_i$), metabolite ($\mathbf{x}$) and input ($\mathbf{u}$) concentrations, respectively, obtained from $p$ experiments denoted by the index $j = {1, 2, ..., p}$.
	\begin{align}\label{eq:theta-eq}
	\theta_k = g_k(\mathbf{v}_i, \mathbf{x}, \mathbf{u}) = \frac{N_k(\mathbf{v}_i, \mathbf{x}, \mathbf{u})}{D_k(\mathbf{v}_i, \mathbf{x}, \mathbf{u})}
	\end{align}

	The identifiability of parameter $\theta_k$, $k = {1, 2, ..., p}$, for flux $v_i$ can be established by determining the value of $D_k(\mathbf{v}_i, \mathbf{x}, \mathbf{u})$ (Figure \ref{fig:ident-flowchart}b): any parameter $\theta_k$ is said to practically identifiable if $D_k(\mathbf{v}_i, \mathbf{x}, \mathbf{u})\neq0$, and practically non-identifiable if $D_k(\mathbf{v}_i, \mathbf{x}, \mathbf{u}) = 0$. Furthermore, the physical properties of the kinetic parameters can be used to distinguish between identifiable and non-identifiable parameter values by designating only parameters with a positive value of $g_k(\mathbf{v}_i, \mathbf{x}, \mathbf{u})$ as identifiable (Figure \ref{fig:ident-flowchart}b). The solution $g_k(\mathbf{v}_i, \mathbf{x}, \mathbf{u})$ in Equation (\ref{eq:theta-eq}) is unique for an identifiable $\theta_k$, and an infinite number of solutions are possible for a non-identifiable $\theta_k$. However, if there are multiple but finite solutions $g_k(\mathbf{v}_i, \mathbf{x}, \mathbf{u})$, then the corresponding parameter $\theta_k$ is locally identifiable.

	\subsection{Numerical parameter identification and estimation}\label{sec:numerical_method}
	The identification methodology described in the previous section can be extended to be used in conjunction with an optimization-based procedure for numerical estimation of parameters from experimental data. The $\mathbb{R}^p$ nonlinear algebraic equations formulated for each flux $v_i$ (Equation \ref{eq:nonlineq}) can be written as shown in Equation (\ref{eq:numerical_opt}) below to be solved as an optimization problem. Minimization of the sum of the absolute error values is used as the objective, and the $\mathbb{R}^p$ nonlinear algebraic equations are provided as inequality constraints.

	\begin{subequations}\label{eq:numerical_opt}
		\begin{align}
		\underset{\theta, \varepsilon}{min} & \sum_{j=1}^{p} \left|\varepsilon_j\right|\\
		& v_{i, j} - f_j(\mathbf{x}_j,\mathbf{\theta}, \mathbf{u}_j) \le \varepsilon_j && j=\{1, 2, ..., p\}\\
		& \theta_l \le \theta_j \le \theta_u && j=\{1, 2, ..., p\}
		\end{align}
	\end{subequations}

	The formulation of the above optimization problem is similar to the least squares problem shown earlier in Equation (\ref{eq:lsqopt}). Notice that the parameter values are bounded above and below by $\theta_u$ and $\theta_l$, respectively. These bounds can be specified to be arbitrarily large to include all possible parameter values.

	\subsection{Degree of identifiability: A quantitative measure of practical identifiability}\label{sec:degree_of_identifiability}
	We express the practical identifiability of kinetic parameters using a simple quantitative term called the degree of identifiability. We describe the degree of identifiability of any single parameter as the percentage of all data combinations (used to test for practical identifiability) that can identify that parameter.

	As an example, if 90\% of all the experimental data combinations used for testing can identify a parameter $\theta_i$, then the degree of identifiability of $\theta_i$ is said to be 0.9 or 90\%. On the other hand, if only 10\% of the combinations can identify another parameter $\theta_j$, then $\theta_j$ has a degree of identifiability of 0.1 or 10\%. Furthermore, we can create a hierarchy of practically identifiable parameters using their degrees of identifiability. In the above instance of the two parameters $\theta_i$ and $\theta_j$ that have degrees of identifiability of 90\% and 10\% respectively, $\theta_i$ is classified to be more identifiable than $\theta_j$ due to its relatively higher degree of identifiability.

	Determining this hierarchy of identifiable parameters can help in distinguishing parameters that can be identified by any type and any combination of experiments from parameters that can be identified by only a select type and combination of experiments. Such a classification can subsequently be used to design minimal sets of experiments that can practically identify all kinetic parameters used to model a metabolic network, going from the least identifiable parameter to the most identifiable parameter.

	\subsection{Kinetic model of gluconeogenesis in \textit{E. coli}}\label{sec:small-model}
	A previously proposed kinetic model \parencite{Kotte2014, Srinivasan2017} for acetate consumption through gluconeogenesis (Figure \ref{fig:network}) is used as a case study to illustrate identifiability analysis for experimental design for parameter estimation in kinetic models of metabolism. The kinetic model is described below.

	\begin{figure}[!tbhp]
		\centering{\includegraphics[width=.3\textwidth,height=.6\textheight,keepaspectratio]{figures/figure5/Figure1_NetworkA}}
		\caption{The previously published small metabolic network for gluconeogenesis used to demonstrate our practical identifiability method for kinetic models of metabolism.}\label{fig:network}
	\end{figure}

	\begin{equation}\label{eq:ode1}
	\frac{d}{dt}pep=v_1-v_2-v_4
	\end{equation}
	\begin{equation}\label{eq:ode2}
	\frac{d}{dt}fdp=v_2-v_3
	\end{equation}
	\begin{equation}\label{eq:ode3}
	\frac{d}{dt}E=v_5 - d E
	\end{equation}
	The kinetic expressions for fluxes $v_1$ through $v_5$ are given below. The consumption of acetate through $v_1$ and conversion of \textit{pep} through $v_2$ are expressed in Equations (\ref{eq:flux1}) and (\ref{eq:flux2}) respectively using Michaelis-Menten kinetics. The acetate flux through $v_1$ is also governed by the quantity of available enzyme E. 
	\begin{equation}\label{eq:flux1}
	v_1 = k_{1}^{cat}E\frac{ac}{ac+K_{1}^{ac}}
	\end{equation}		
	The model for flux $v_1$ of the small network (Figure \ref{fig:network}), uses the concentration of the enzyme E as a variable (Equation \ref{eq:flux1}). Since we assume that steady state experimental information is only available for metabolite concentrations and fluxes, and not for enzymes (again the details on relaxing this assumption are discussed later), the expression in Equation (\ref{eq:flux1}) for $v_1$ cannot be used for identifying parameters $k_1^{cat}$ and $K_1^{ac}$. So, we modify the Michaelis-Menten kinetic rate law expression to eliminate the enzyme concentration E as a variable in Equation (\ref{eq:flux1a}). Consequently $k_1^{cat}$ is replaced by $V_1^{max}$ as a parameter to describe $v_1$. The corresponding enzyme binding constant is denoted as $K_1^{ac} (ne)$ to distinguish it from the enzyme binding constant calculated in the presence of measured enzyme concentration data.
	\begin{align}\label{eq:flux1a}
	v_1 = V_1^{max}\frac{ac}{ac+K_{1}^{ac}(ne)}
	\end{align}		
	We choose the expression for flux $v_1$ given in Equation (\ref{eq:flux1a}) to demonstrate our method for practical identifiability. 	
	\begin{equation}\label{eq:flux2}
	v_2 = V_{2}^{max}\frac{pep}{pep+K_{2}^{pep}}
	\end{equation}
	\begin{equation}\label{eq:flux3}
	v_3 = V_{3}^{max}\frac{\tilde{fdp}\left(1+\tilde{fdp}\right)^3}{\left(1+\tilde{fdp}\right)^4+L_3\left(1+\frac{pep}{K_{3}^{pep}}\right)^{-4}}
	\end{equation}
	The allosterically regulated flux $v_3$ for the consumption of \textit{fdp} is expressed in Equation (\ref{eq:flux3}) using the Monod-Wyman-Changeux (MWC) model for allosterically regulated enzymes, where $\tilde{fdp}$ refers to the ratio of \textit{fdp} with respect to its allosteric binding constant $K_{3}^{fdp}$.

	The practically identifiability of parameters of a given flux are determined by solving a system of nonlinear algebraic equations using a computer algebra system (Section \ref{sec:ident}). We find that the nonlinearity of the MWC kinetic rate law used to model the allosteric regulation of $v_3$ makes it computationally intractable for determining the closed form expressions of the three parameters $V_3^{max}$, $K_3^{fdp}$ and $K_3^{pep}$ using a computer algebra system (Mathematica or SymPy in Python). In order to overcome this computational obstacle, we model the reaction rate for $v_3$ using the convenience kinetic rate law formulation \parencite{Liebermeister2006}. The corresponding expression obtained for $v_3$ is given below (Equation \ref{eq:flux3_convkin}). 	
	\begin{align}\label{eq:flux3_convkin}
	v_3 = V_3^{max}\left(\frac{1}{1 + \frac{K_3^{pep}}{pep}}\right)\left(\frac{\frac{fdp}{K_3^{fdp}}}{1 + \frac{fdp}{K_3^{fdp}}}\right)
	\end{align}

	The flux $v_4$ for the export of \textit{pep} is expressed as a linear equation dependent on $pep$ in Equation (\ref{eq:flux4}).
	\begin{equation}\label{eq:flux4}
	v_4 = V_{4}^{max}.pep
	\end{equation}		
	The production of enzyme E is represented by flux $v_5$. The inhibition of this flux by \textit{fdp} is modeled using Hill kinetics, where $K_e^{fdp}$ represents the Hill binding constant for the inhibiting metabolite \textit{fdp}, $n_e$ is the Hill exponent, and $V_e^{max}$ is the maximum reaction rate for $v_5$.
	\begin{align}\label{eq:flux5}
	v_5 = V_e^{max}\left(\frac{1}{1+\left(\frac{fdp}{K_{e}^{fdp}}\right)^{n_e}}\right)
	\end{align}

	\subsection{Experimental design through practical parameter identification}\label{sec:experimental_design}		
	Not all metabolite concentrations and fluxes in the model (Equation \ref{eq:kinstoich}) change for any random experiment. This makes unambiguous estimation of parameters impossible, either due to the inherent correlation between changes in different concentrations or fluxes, or due to the homeostasis of the concentrations and fluxes under the chosen experimental conditions \parencite{Heijnen2013}. In such scenarios, the need to design experiments to effect a change in, and discriminate between changes in different concentrations/fluxes becomes necessary.

	Following the methodology described in Section \ref{sec:ident}, and demonstrated in Section \ref{sec:example} for a single flux using data from a combination of two different experiments, all distinct combinations of data sets obtained from experiments described in Section \ref{sec:experiments} of the Supplementary Information can be tested for their ability to practically identify any of the fluxes in the small metabolic network. This step would determine the degree of identifiability (defined in Section \ref{sec:degree_of_identifiability}) of each parameter in each flux in the model, and help distinguish experiment combinations that contribute to identifiability from combinations that do not practically identify any parameter in the model (Figure \ref{fig:ident-flowchart}b). In doing so, it is possible to obtain a minimal and informative collection of experiments that can be performed to identify as many model parameters as possible (Figure \ref{fig:ident-design}). Consequently, the set of experiments can be used to estimate all the identifiable parameters in the model. This is formally explained below.

	The identifiability of each parameter based on each experiment with index $j = {1, 2, ..., n_E}$ is established based on the methodology described in Section \ref{sec:ident} (Figure \ref{fig:ident-flowchart}b), and demonstrated in Section \ref{sec:example}. Subsequently, for any flux $v_i$, and for any combination of $p$ experimental data sets, if the experimental concentrations and fluxes ($\mathbf{x}_j$ and $\mathbf{v}_j$, respectively, where $j = {1, 2,..., p}$) do not satisfy the condition for identifiability for any parameter $\theta_k$ in $\theta\in\mathbb{R}^p$ (Figure \ref{fig:ident-flowchart}b), then at least one of the $p$ experiments needs to be changed to make parameter $\theta_k$ identifiable. Consequently, the corresponding experiment cannot be used for estimating parameter $\theta_k$, and needs to be discarded from the set of all necessary experiments. Furthermore, another experiment from $j = {1, ..., n_E}$ needs to be selected to replace the discarded experiment such that parameter $\theta_k$ is identifiable. This process has to be repeated until all parameters in $\theta\in\mathbb{R}^p$ are identifiable for flux $v_i$. In doing so, we can arrive at a set of $p$ experiments that will always result in practically identifiable parameters for flux $v_i$. Note that if none of the $n_E$ pre-selected experiments satisfy the identifiability condition, then we can design an $(n_E+1)^{th}$ experiment that can replace one of the experiments that causes practical non-identifiability. This analysis can be performed for each flux in a metabolic network independent of all the other fluxes, making it theoretically scalable even to genome-scale models of metabolism.

	\section{Results}
	First, in Section \ref{sec:example}, we demonstrate the use of the methodology that we described in Section 2.1 to practically identify parameters in flux $v_1$ of the small gluconeogenic network (Figure \ref{fig:network}) model given in Section \ref{sec:small-model}. 
	We discuss the ability of the proposed methodology to determine the structural identifiability of parameters modeling $v_1$, $v_3$ and $v_5$ in Section \ref{sec:proof}. 	
	In Section \ref{sec:initial_analysis} that follows, we show how the demonstrated methodology is capable of practically identifying and estimating parameters for fluxes $v_1$, $v_2$, $v_3$ and $v_5$ using steady state flux values and metabolite concentrations. 
	The various ways in which this information can be used for designing experiments to generate data that can facilitate estimation of identifiable parameters are discussed in Section \ref{sec:design}.
	The contribution of the uncertainty in the data arising from either the differences between in vivo and in vitro kinetics, or the noise present in experimentally measured quantities towards identifying parameters in enzyme kinetic models is discussed finally in Section \ref{sec:uncertainty}.

	\subsection{Identifying parameters in kinetic models of metabolism: an example}\label{sec:example}	
	In this section, we illustrate the proposed methodology step by step to identify parameters of flux $v_1$ in the small metabolic network (Figure \ref{fig:network} and Section \ref{sec:small-model}). We choose the expression for flux $v_1$ given in Equation (\ref{eq:flux1a}) for this demonstration.

	Since $\theta = \{V_1^{max}, K_1^{ac} (ne)\} \in \mathbb{R}^2$ for $v_1$, as mentioned in Section \ref{sec:experiments}, we need steady state concentration and flux measurements from at least two different experiments. So, from the $n_E = 21$ different experiments described in Section \ref{sec:experiments} and Table \ref{tab:pval}, we can choose multiple combinations of $p = 2$ experiments to satisfy the data requirements for identifying $v_1$ i.e., in Equation (\ref{eq:nonlineq}) $j = \{1, 2\}$. We label the available concentrations and fluxes as ${ac}^{(j)}$ and ${v_1}^{(j)}$, respectively. Then, the nonlinear algebraic equations shown in Equation (\ref{eq:nonlineq}) can be formulated for $v_1$ as:
	\begin{align*}%\label{eq:nonlin-flux2}
	{v_1}^{(j)} = V_{1}^{max}\frac{ac^{(j)}}{ac^{(j)}+K_{1}^{ac}(ne)} &&  j=\{1, 2\}
	\end{align*}

	Solving this simultaneous system of equations in $\mathbb{R}^2$ using Mathematica (Wolfram Research, USA), a computer algebra system, we get $p=2$ nonlinear algebraic expressions for parameters $V_1^{max}$ (Equation \ref{eq:v1_par}a) and $K_1^{ac}(ne)$ (Equation \ref{eq:v1_par}b). These expressions have the form shown in Equation (\ref{eq:theta-eq}).
	\begin{subequations}\label{eq:v1_par}
		\begin{align}		
		\theta_1 &= V_1^{max} = \frac{v_1^{(1)}v_1^{(2)}(ac^{(1)}-ac^{(2)})}{v_1^{(2)}ac^{(1)}-v_1^{(1)}ac^{(2)}}\\
		\theta_2 &= K_1^{ac}(ne) = \frac{ac^{(1)}ac^{(2)}(v_1^{(1)}-v_1^{(2)})}{v_1^{(2)}ac^{(1)}-v_1^{(1)}ac^{(2)}}
		\end{align}
	\end{subequations}
	To test the practical identifiability of the parameters in Equation \ref{eq:v1_par}, we substitute any suitable in silico experimental data and determine the value of the denominator of the right hand side expression. Since the enzyme binding constant ($K_1^{ac}(ne)$) and the maximum reaction rate ($V_1^{max}$) cannot be negative, we can further constrain the criteria for identifiability for both these parameters by saying that the evaluated expressions in Equation (\ref{eq:v1_par}) should be positive (Figure \ref{fig:ident-flowchart}b). The parameter values that are obtained for $V_1^{max}$ and $K_1^{ac} (ne)$ by substituting in silico steady state experimental data are shown in Supplementary Figure \ref{fig:v1_v1max_ck_values}. Due to the numerous possible parameter values seen in Supplementary Figure \ref{fig:v1_v1max_ck_values}, we can conclude that both $V_1^{max}$ and $K_1^{ac}$ are practically non-identifiable.

	We can also apply the proposed methodology to practically identify parameters in $v_1$ under the assumption that the protein concentration for the enzyme \textit{E} is also available, in addition to the measured metabolite concentrations and fluxes. In doing so, we get two expressions similar to the one shown in Equation (\ref{eq:v1_par}) for $k_1^{cat}$ and $K_1^{ac}$. Here, the value of $V_1^{max}$ in Equation (\ref{eq:flux1a}) is substituted with $V_1^{max} = k_1^{cat}E$ instead. The corresponding identifiability expressions for $k_1^{cat}$ and $K_1^{ac}$ are given in Equation (\ref{eq:v1_k1cat}).	
	\begin{subequations}\label{eq:v1_k1cat}
		\begin{align}
		k_1^{cat} = \frac{ v_1^{(1)}v_1^{(2)} \left( ac^{(1)}- ac^{(2)}\right)}{v_1^{(2)}ac^{(1)}E^{(1)} - v_1^{(1)}ac^{(2)}E^{(2)}}\\
		K_1^{ac} = \frac{ac^{(1)}ac^{(2)}\left(v_1^{(1)}E^{(2)} - v_1^{(2)}E^{(1)}\right)}{v_1^{(2)}ac^{(1)}E^{(1)} - v_1^{(1)}ac^{(2)}E^{(2)}}
		\end{align}
	\end{subequations}	
	We show the parameter value for $k_1^{cat}$ and $K_1^{ac}$ that are obtained through the practical identifiability analysis in Figure \ref{fig:ident_values}a when in silico experimental data is substituted in Equation (\ref{eq:v1_k1cat}). Through Equation (\ref{eq:v1_k1cat}) and Figure \ref{fig:ident_values}a we are able to show that the uncertainty in the parameter estimates (Supplementary Figure \ref{fig:v1_v1max_ck_values}) can be resolved through the incorporation of the available enzyme concentrations. Thus, having more experimental information can help resolve practical identifiability.

	\begin{figure}[!tbhp]
		\centering{\includegraphics[width=.8\textwidth,height=.6\textheight,keepaspectratio]{figures/figure6/v1_k1cat_v2_v5_2_v3_ck_1_parameter_values}}
		\caption{Distribution of predicted parameter values when performing practical identifiability analysis using closed-form solutions for each parameter in flux a) $v_1$, b) $v_2$, c) $v_5$, and d)$v_3$. For $v_1$, we have assumed that enzyme concentration is available and have accordingly identified and estimated $k_1^{cat}$, as opposed to $V_1^{max}$. The parameter values for only the second root of $K_e^{fdp}$ in $v_5$ ($K_e^{fdp}(2)$) is shown, since $K_e^{fdp}(1)$ is not estimated by any combination of two experiments, and $V_e^{max}$ is estimated by all combinations. Only one of the two roots for $v_3$ is shown in panel d. The estimated data for the second root has a similar distribution to that of the first root and is shown in the Supplementary Information. Data is generated using the Convenience Kinetic model for allosteric regulation for $v_3$.}\label{fig:ident_values}
	\end{figure}

	In the following section we present results from the identifiability analysis of fluxes $v_2$, $v_3$ and $v_5$ in the small metabolic network (Figure \ref{fig:network}), using the methodology (Figure \ref{fig:ident-flowchart}) that we have demonstrated above for $v_1$.

	\subsection{Establishing Structural identifiability of parameters based on closed-form solutions}\label{sec:proof}
	For the proposed methodology (Figure \ref{fig:ident-flowchart}) to work, it should be possible to obtain closed form solutions for each parameter in the enzyme kinetic model for each flux as shown in Equation (\ref{eq:theta-eq}). Since the ability to obtain closed-form solutions for each parameter is dependent on the model structure, any parameter that has non-unique closed-form solutions can be called a structurally non-identifiable parameter. However, if the number of solutions that the parameter has are finite, then the parameter is only locally structurally identifiable.

	We demonstrated the structural identifiability of parameters modeling $v_1$ in Section \ref{sec:example}. We have shown that the parameters have only one unique closed-form solution, and accordingly are structurally identifiable. Since $v_2$ is also expressed using the Michaelis-Menten model, just like $v_1$, we find that the parameters ($V_2^{max}$ and $K_2^{pep}$) are also structurally identifiable. The closed-form expressions for these parameters are similar to the ones shown in Equation \ref{eq:v1_par}, with \textit{ac} replaced by \textit{pep}, and $v_1$ replaced by $v_2$.

	However, we find that the parameters used to model $v_3$ using the Convenience kinetics rate law, and $v_5$ using the Hill kinetic rate law are not structurally identifiable as they are given in Section \ref{sec:small-model}.

	First, for $v_3$, we find that the parameters $V_3^{max}$, $K_3^{fdp}$ and $K_3^{pep}$ have two different closed-form solutions. Thus, based on the presence of non-unique but finite number of possible solutions for these parameters we can classify $v_3$ as a locally structurally identifiable flux. In order to alleviate local structural identifiability, we reduced the dimension of the parameter space for $v_3$. Originally, $\theta \in \mathbb{R}^3$ for $v_3$. By reducing the dimension of $\theta$ to $\mathbb{R}^2$, we were able to obtain a structurally identifiable model for $v_3$. To reduce the dimension of the parameter space for $v_3$, we fix either $K_3^{fdp}$ or $K_3^{pep}$ as a known quantity, and identify the other unfixed parameter along with $V_3^{max}$. This results in unique closed-form expressions for both $V_3^{max}$ and the other unfixed parameter ($K_3^{pep}$ or $K_3^{fdp}$).

	%We would like to point out that we used the Convenience kinetics rate law model for $v_3$ due to the inability of the computer algebra system to generate closed-form expressions for parameters in the Monod-Wyman-Chageaux model (Section \ref{sec:small-model}). However, we find that the Convenience kinetics model is not structurally identifiable in its current form.

	While $v_3$ is an allosterically regulated metabolic flux, $v_5$ describes a transcription/translation reaction using Hill kinetics. We apply our proposed methodology to identify parameters modeling $v_5$ using only the available experimental data on the metabolite concentrations and the fluxes within the metabolic network. We could not obtain closed form solutions for parameters $V_e^{max}$, $K_e^{fdp}$ and $n_e$ in $v_5$ using the computer algebra system. So, instead of changing the model as we did for $v_3$ (see Section \ref{sec:small-model}), we resorted to reducing the dimension of the parameter space by fixing one of the three parameters, the Hill coefficient $n_e$. We illustrated the consequence of reducing the dimension of the parameter space of the Convenience kinetic model for $v_3$ earlier. With a fixed and known $n_e$, $K_e^{fdp}$ has two possible closed-form solutions, which make it a locally structurally identifiable parameter. On the other hand, $V_e^{max}$ has only one unique closed-form solution, and therefore is structurally identifiable.

	We have now established conditions for structural identifiability of all the major fluxes in the small metabolic network (Figure \ref{fig:network}). We have shown how our proposed methodology can be used to establish conditions for structural identifiability using steady state information on the model variables. We next discuss the practical identifiability of the parameters in $v_2$, $v_3$ and $v_5$ whose parameters are structurally identifiable only under certain conditions.

	\subsection{Relationship between structural and practical parameter identifiability}\label{sec:initial_analysis}
	We mention in Section \ref{sec:ident_def} that, by definition, unique parameter values based on the model structure are possible for any structurally identifiable parameter. Together with this definition for structural identifiability, we also introduced the concept of practical parameter identifiability. To recall, we mentioned that  it should be possible to estimate unique parameter values based on all available experimental data for any practically identifiable parameter.

	As shown in Figure \ref{fig:ident-flowchart} and illustrated for $v_1$ in Section \ref{sec:example}, to determine the practical identifiability of parameters we test for the existence of a non-zero denominator of the closed-form expressions of the parameters. We also reduce the possible space within which a parameter could be practically identifiable by checking for the physiological feasibility of the parameter values that are obtained through this analysis (Figure \ref{fig:ident-flowchart}b). If the resulting parameter values obtained from various combinations of experimental data for each closed-form expression are unique, then the parameter is practically identifiable. However, if a non-unique number of parameter values are possible from multiple combinations of experimental steady state data, then the parameter is said to be practically non-identifiable. In conjunction with the conditions for structural identifiability demonstrated earlier in Section \ref{sec:proof}, if the parameter has only one unique closed-form expression, and its value is also unique, then the parameter is both structurally and practically identifiable. If either of these conditions are not satisfied, the parameters can be either locally structurally or practically identifiable or non-identifiable.

	Accordingly, both $v_1$ and $v_2$ are not only structurally identifiable due to the presence of unique closed-form expressions for their parameters, they are also practically identifiable because the parameters in the respective models possess unique values based on distinct combinations of experimental data (Figure \ref{fig:ident_values}a and b).

	Regarding $v_5$, we showed earlier in Section \ref{sec:proof} that the identifiability of $v_5$ can be analyzed only when the Hill coefficient $n_e$ is held constant. So, in subsequent discussions, the dimension of the $v_5$ parameter space is kept at $\mathbb{R}^2$ by fixing the value of $n_e$. Under these conditions, we find that the structurally identifiable parameter $V_e^{max}$ is also practically identifiable, i.e., it has only one unique value based on all available in silico experimental data (Figure \ref{fig:ident_values}c). However, recall that unlike $V_e^{max}$, $K_e^{fdp}$ is only locally structurally identifiable as it has two possible closed-form expressions. Nonetheless, despite its local structural identifiability, we find that the $K_e^{fdp}$ is also practically identifiable, like $V_e^{max}$, with only one unique parameter value (Figure \ref{fig:ident_values}c).

	We find that the practical identifiability of $v_5$, despite the local structural identifiability of one of its parameters, is due to the enforcement of the physiological relevance criteria on the parameters i.e., only one of the two closed-form expressions for $K_e^{fdp}$ is physiologically relevant. The other solution always acquires a negative value that has no physiological meaning. Thus, by reducing the practically identifiable space of parameters, we have shown that our methodology can establish global practical identifiability even when the parameters are only locally structurally identifiable.

	Similar to $K_e^{fdp}$ in $v_5$, we also explained the local structural identifiability of $V_3^{max}$, $K_3^{fdp}$ and $K_3^{pep}$ modeling $v_3$ in Section \ref{sec:proof}. These parameters have two possible closed-form expressions. In Figure \ref{fig:ident_values}d we show the numerical values for one of the two possible closed-form expressions for $V_3^{max}$, $K_3^{fdp}$ and $K_3^{pep}$. The numerical values for the second closed-form expressions of the three parameters is presented in Supplementary Figure, and they also have a similar distribution. Based on the prior definition for practically identifiable parameters, the numerous possible values that the three parameters can acquire (Figure \ref{fig:ident_values}d) leads us to conclude that the parameters in $v_3$ are practically non-identifiable when they are only structurally locally identifiable. Also, unlike $v_5$, which is practically identifiable in the presence of local structural identifiability, parameters for $v_3$ are practically non-identifiable even after the reduction in the practically identifiable parameter space realized using the physiological relevance condition (Figure \ref{fig:ident-flowchart}b).

	\begin{figure}[!tbhp]
		\centering{\includegraphics[width=0.8\textwidth,height=0.3\textheight, keepaspectratio]{figures/figure6/v3_var_1_2_ck_parameter_value}}
		\caption{Distribution of predicted parameter values when performing practical identifiability analysis using closed-form solutions for each parameter in flux $v_3$. The globally identifiable parameter values of a)  $V_3^{max}$ and $K_3^{fdp}$ when $K_3^{pep}$ is held constant, and b) $V_3^{max}$ and $K_3^{pep}$ when $K_3^{fdp}$ is held constant.}\label{fig:v3_var_ck_values}
	\end{figure}

	%So, the observation of practical non-identifiability for $v_3$ in the presence of local structural identifiability is in contrast to what we observed for $v_5$ earlier. 
	However, we find that $v_3$ is practically identifiable when its parameters are also structurally identifiable (Figure \ref{fig:v3_var_ck_values}). Earlier in Section \ref{sec:proof} we had mentioned that $V_3^{max}$ and $K_3^{fdp}$ are structurally identifiable only when $K_3^{pep}$ is fixed, and $V_3^{max}$ and $K_3^{pep}$ are structurally identifiable when $K_3^{fdp}$ is fixed. Under these scenarios we find the structurally identifiable parameters to also be practically identifiable (Figure \ref{fig:v3_var_ck_values}).

	In conjunction with the practical identifiability of $v_5$ established earlier, we see that it is possible to delineate between structural and practical identifiability of parameters in kinetic models of metabolism only under certain conditions, and not in others.

	\subsection{A priori experimental design through practical parameter identification}\label{sec:design}	
	The analysis of parameter practical identifiability can be used to gather information on the type of experiments that can provide useful data for parameter estimation. For instance, during practical identifiability analysis, if either the denominator of the closed-form expression is zero, or if the parameter values that are obtained are not physiologically feasible (Figure \ref{fig:ident-flowchart}b), then the experimental data set concerned is said to be incapable of practically identifying that said parameter. Consequently, the data from that combination of experiments is considered non-informative. When this analysis is repeated for multiple combinations of steady state data from the 21 different in silico experiments, we can determine the number of different experimental data combinations that can practically identify each parameter in each flux (Figure \ref{fig:ident}).

	\begin{figure}[!tbhp]
		\centering{\includegraphics[width=0.6\textwidth,height=0.4\textheight,keepaspectratio]{figures/figure1/v1_k1cat_v2_v5_2_ck_ident}}
		\caption{The number of data combination from 21 different in silico experiments that can practically identify each parameter in fluxes a) $v_1$, b) $v_2$, and c) $v_5$ when there is no noise in the input experimental data. The percentage of total combinations of experimental data used for analysis (240 for $v_1$ and $v_2$, and 421 for $v_5$) that can identify each parameter is also specified. $v_1$, $v_2$ and $v_5$ require data from two experiments for analysis. The contribution of different experiment type towards identifying each parameter is shown in the spider plots.}\label{fig:ident}
	\end{figure}

	As described in Section \ref{sec:degree_of_identifiability}, the information on the number of experimental data sets that can practically identify each parameter can be used to determine the degree of identifiability of the corresponding parameters. Subsequently, this information can be used to classify parameters based on their ease of identifiability.

	In Figure \ref{fig:ident} we show the number of experimental data combinations that are capable of identifying each parameter, and consequently, the degree of identifiability of each parameter (percentage experimental data combinations that are capable of identifying each parameter) in flux $v_1$ (Figure \ref{fig:ident}a), $v_2$ (Figure \ref{fig:ident}b) and $v_5$ (Figure \ref{fig:ident}c). The degree of identifiability for $V_1^{max}$ and $K_1^{ac}(ne)$ in $v_1$ is shown in Supplementary Figures \ref{fig:v1_v1max_ck_ident}, and the degree of identifiability for both closed-form expressions of the three parameters of $v_3$ are shown in Supplementary Figure \ref{fig:v3_1_ck_ident}. The degree of identifiability of each parameter is also given in these figures as percentages.

	It is important to recall that both $v_1$ (Supplementary Figure \ref{fig:v1_v1max_ck_values}) and $v_3$ (Figure \ref{fig:ident_values}d) are not practically identifiable (Section \ref{sec:initial_analysis}). While $v_1$ becomes practically identifiable (Figure \ref{fig:ident_values}a) if enzyme concentrations are utilized to alleviate uncertainties in the enzyme turn over rates ($k_1^{cat}$), $v_3$ becomes practically identifiable only when the dimension of the parameter space is reduced (Figure \ref{fig:v3_var_ck_values}). In these cases the degree of identifiability (Supplementary Figure \ref{fig:v1_v1max_ck_ident} for $v_1$ and Supplementary Figure \ref{fig:v3_1_ck_ident} for $v_3$) refers to the number of experimental data sets that can determine physiologically relevant values for the corresponding parameters.

	Based on their degrees of identifiability, we see that the maximum reaction rates ($V_i^{max}$) are more or similarly identifiable in comparison to the corresponding enzyme binding ($K_i$) constants or the activation/inhibition constants, in the respective reaction rate law models (Figure \ref{fig:ident}, Supplementary Figures \ref{fig:v1_v1max_ck_ident} and \ref{fig:v3_1_ck_ident}). We make this observation despite the fact that the four fluxes are modeled using three different enzyme kinetic rate laws: $v_1$ and $v_2$ are modeled using the Michaelis-Menten rate law, $v_3$ is modeled using the Convenience kinetic rate law and $v_5$ is modeled as a Hill equation with inhibition. As mentioned earlier, for $v_1$ (Supplementary Figure \ref{fig:v1_v1max_ck_ident}) and $v_3$ (Supplementary Figure \ref{fig:v3_1_ck_ident}) we have shown that a greater number of experimental data sets can predict physiologically relevant or non-zero positive values for $V_1^{max}$ and $V_3^{max}$ than for $K_1^{ac}(ne)$ and $K_3^{fdp}$ or $K_3^{pep}$, respectively.

	The degree of identifiability of $v_3$, when its parameters are structurally and practically identifiable (Sections \ref{sec:proof} and \ref{sec:initial_analysis}) are shown in Supplementary Figure \ref{fig:v3_var_1_2_ck_ident}. Accordingly, we find that with the exception of $V_1^{max}$ (Supplementary Figure \ref{fig:v1_v1max_ck_ident}) and $k_1^{cat}$ in $v_1$ (Figure \ref{fig:ident}a), all data sets used to test practical identifiability can determine unique values for parameters when the corresponding parameter is structurally identifiable.

	We can attribute the difference in the degree of identifiability between $v_1$ (Figure \ref{fig:ident}a and Supplementary Figure \ref{fig:v1_v1max_ck_ident}) and the other fluxes ($v_2$, $v_3$ and $v_5$) to the ability of data from different combinations of experiments to satisfy the conditions for practical identifiability of that parameter, that can be determined a priori. In systems identification terminology, data requirements for parameter identification can be tied to selecting experiments that are persistently excitable for the flux being identified. Any input signal should be rich or informative enough to guarantee full excitement of the dynamics of the system \parencite{Ljung1994}. Only information obtained from such changes in the input can be used to completely identify the system over its entire dynamic range. So, the ability of data from a combination of different experiments to practically identify parameters of a given flux is governed by the ability of the experiment to generate distinct measured concentrations and fluxes that will satisfy the identifiability conditions.

	In turn, the degree of identifiability of parameters and the informativeness of the corresponding experiments used to identify them can be explained by the position of the flux in the metabolic network. The position of any given flux in the metabolic network determines the specific experiment that is persistently excitable enough to identify the parameters of that flux. This dependency can be further elucidated using $v_1$ and $v_2$ as examples.

	We know from Equation (\ref{eq:v1_k1cat}) and Section \ref{sec:example} that for a combination of any two experiments to be capable of identifying $v_1$, the experiments must generate data that have distinct acetate concentrations, \textit{E} and $v_1$. We also know, based on our knowledge of the Michaelis-Menten kinetic rate law that changes in the substrate concentration of a reaction can bring about a nonlinear change in the value of the corresponding reaction rate. So, in this instance, since the substrate is an input variable to the model, and $v_1$ is the corresponding uptake flux and \textit{E} is a system variable, the substrate can be easily perturbed to create persistently excitable experiments to identify parameters in $v_1$. We see the consequence of this requirement in the degree of identifiability of $k_1^{cat}$ and $K_1^{ac}$ (Figure \ref{fig:ident}a). We can generalize this observation for the identification of all uptake fluxes in all metabolic networks, i.e., at a minimum, a change in the input substrate concentration may be necessary for an informative experiment to identify the uptake flux parameters.

	Similarly, the identification of parameters for $v_2$ (Figure \ref{fig:ident}b) requires that persistently excitable experiments distinguish between values of both $v_2$ as well as \textit{pep}. However, since both of these are system outputs, satisfaction of this condition cannot be guaranteed without an analysis of the dynamics of the metabolic network, and how changes in the input (acetate) bring about changes in the two requisite output quantities. Previous dynamical analysis of the network (Figure \ref{fig:network}) has already established the existence of a functional relationship between \textit{pep} and $v_2$, and the input acetate concentration and the levels of expression of the different enzymes within the network \parencite{Srinivasan2017}. The 100\% degree of identifiability seen for $v_2$ (Figure \ref{fig:ident}b) confirms the theoretical possibility for any type of perturbation experiment to be persistently excitable to identify $v_2$. Overall, this analysis informs us that the degree of identifiability and consequently, the type of experiments needed to identify different parameters varies widely depending on the position of the flux with respect to the inputs and the outputs of the metabolic network, as well as the various regulatory interactions present within the network (e.g., effect of \textit{pep} on $v_3$, or the effect of \textit{fdp} on $v_5$ and consequently on $v_1$ in Figure \ref{fig:network}).

	\begin{figure}[!tbhp]
		\centering{\includegraphics[width=1.0\textwidth,height=0.5\textheight,keepaspectratio]{figures/figure2/v1_k1cat_v2_v5_2_ck_exp_info}}
		\caption{Frequency of each type of perturbation experiment in data sets that can identify parameters in fluxes a) $v_1$, b) $v_2$ and c)$v_5$. The frequencies are represented as a percentage of the total number of experiments present within all the data sets that can practically identify the parameters. The frequency of experiments that occur as the first experiment in a combination of two experiments are shown as a blue surface, and the frequency of experiments that occur as the second experiment in the combination are shown as a green surface.}\label{fig:expinfo}
	\end{figure}

	From the above example we can summarize that identification of individual fluxes within a metabolic network necessitates a careful consideration of experiments such that the data acquired can satisfy conditions for practical identifiability for all parameters modeling a flux, and subsequently, all fluxes within a network \parencite{Heijnen2013}. To facilitate the design of experiments based on their ability to satisfy requirements for practical identifiability of parameters, we determine the occurrence of each type of steady state perturbation experiment within combinations that can practically identify each parameter (Figure \ref{fig:expinfo}, Supplementary Figures \ref{fig:v1_v1max_ck_ident},  \ref{fig:v3_1_ck_ident} and \ref{fig:v3_var_1_2_ck_ident}). So, with our proposed methodology it is possible to identify the types of perturbation experiments that would be informative for identifying each parameter in each flux with steady state concentration and flux data.

	As mentioned in Supplementary Section \ref{sec:experiments}, we use experimental data from five different types of experiments to test the practical identifiability of parameters in the model (Supplementary Table \ref{tab:pval}). In Figures \ref{fig:expinfo}a, \ref{fig:expinfo}b and \ref{fig:expinfo}c, the contribution from different experiment types for identifying parameters in $v_1$, $v_2$ and $v_5$ are respectively shown as spider plots.

	The contribution of experiments that involve changes in the acetate concentrations, which consequently bring about changes in the value of $v_1$, contribute to a significant part (> 50\%) of the identifiable experimental data combinations for $v_1$ in comparison to the other types of experiments (Figure \ref{fig:expinfo}a and Supplementary Figure \ref{fig:v1_v1max_ck_ident}). This is in agreement with the condition for identifiability that we discussed earlier (Equations \ref{eq:v1_k1cat} and \ref{eq:v1_par}). Since less than 50\% of all data combinations can satisfy these requirements, and can consequently identify $v_1$ (Figure \ref{fig:ident}a), we also say that identifiability analysis is crucial to determine the minimum number of experiments, along with the nature of experiments that can help identify parameters for $v_1$.

	With $v_2$, we see that the enzyme perturbations as well as the acetate perturbation experiments have similar contributions towards datasets that can identify $v_2$ (Figure \ref{fig:expinfo}b). This also supports our arguments made earlier with regards to the identifiability conditions for $v_2$, and the reasons for the difference in the type of experiments that are informative between $v_1$ and $v_2$. Accordingly, we find that in comparison to selecting experiments to identify $v_1$, there is very little restriction on the types of experiments that are informative to identify $v_2$.

	We can also extend these observations to justify the observed contribution of experiments towards identifying parameters for $v_5$ (Figure \ref{fig:expinfo}c), or determining physiologically relevant parameter values for structurally locally identifiable parameters of $v_3$ (Supplementary Figures \ref{fig:v3_1_ck_ident} and \ref{fig:v3_var_1_2_ck_ident}).

	In all of the above scenarios for $v_1$ (Figure \ref{fig:expinfo}a), $v_2$ (Figure \ref{fig:expinfo}b) and $v_3$ (Supplementary Figures \ref{fig:v3_1_ck_ident} and \ref{fig:v3_var_1_2_ck_ident}), the distribution of experiment types between the two ($v_1$ and $v_2$) or three ($v_3$) required experiments is quite similar. Hence, the green/blue/yellow surfaces in the spider plots are superimposed upon each other. This is also seen for experiments identifying $V_e^{max}$ in $v_5$ (Figure \ref{fig:expinfo}c). However, this is not the case for $K_e^{fdp}$ in $v_5$ (Figure \ref{fig:expinfo}c). We see that when two experiments are required to identify $K_e^{fdp}$, the choice of the first experiment has a bearing on the choice of the second experiment, and vice-versa, so that data with enough information is available for the identification of $K_e^{fdp}$. Also, since $V_e^{max}$ is globally identifiable using any experiment type (Figures \ref{fig:ident}c and \ref{fig:expinfo}c), the choice and number of experiments required to identify $v_5$ completely hinges upon the identifiability of $K_e^{fdp}$ from the chosen experiments.

	So, now we have established a hierarchy of parameters based on their identifiability. Parameters for $v_2$ and $V_e^{max}$ in $v_5$ that are most identifiable are at the top of the hierarchy, while $v_1$ and $K_e^{fdp}$ in $v_5$ fall at the bottom of the list as the experiments needed for their identification require careful consideration. When $v_3$ is structurally identifiable, their parameters also do not require any experimental design considerations.

	\subsection{Parameter non-identifiability due to uncertainty in Experimental Data}\label{sec:uncertainty}
	In all the aforementioned scenarios, the kinetic rate law from which data is derived is known and same as the model for which parameters are estimated. However, in reality, the kinetic rate law based on which metabolic networks function and from which in vivo experimental data is extracted is mostly unknown. The rate laws are primarily inferred through the parameter estimation procedure. This is one of the motivations for the development of approximate kinetic rate law models \parencite{Heijnen2013,Smallbone2007,Berthoumieux2013}.
	So, there is a need to see if the methodology that we have developed here is capable of handling the uncertainty that arises due to the mismatch between the model and the data used to identify and estimate the parameters in the model.

	The scope within which we have defined the model (Section \ref{sec:small-model}) makes such an analysis possible by changing the enzyme kinetic rate law used to describe $v_3$. Note that the original description \parencite{Kotte2014,Srinivasan2017} of the network (Figure \ref{fig:network}) uses the Monod-Wyman-Chageaux (MWC) model to describe the flux through $v_3$. Whereas, so far we have used a Convenience kinetic rate law description for both data generation as well as identifiability analysis. To determine the ability of our methodology to handle the in vivo-in vitro model uncertainty, we use the MWC model description to generate the in silico experimental data. This data will then be used to identify parameters in all the fluxes, including $v_3$ that is described by the Convenience kinetic model.

	First, we find that the spread in the estimated Convenience kinetics parameter values, when $v_3$ is only locally structurally identifiable, is much larger than when there is mismatch between the model generating the data and the model that is being identified (Supplementary Figure). A more important observation is that even when the parameters are structurally identifiable in $v_3$ (achieved by assuming either $K_3^{fdp}$ or $K_3^{pep}$ as a known constant), they can at most only be locally practically identifiable. This is shown by the spread in the estimate values of the structurally identifiable parameters when steady state data based on the MWC model is used in Supplementary Figure.

	Second, note that the dynamics of the network as represented by an MWC model for $v_3$ are different from the dynamic characteristics expressed when a Convenience kinetics model is used instead to describe $v_3$. Thus, this can bring about a change in the steady state concentrations and fluxes observed for the various in silico experiments listed in Supplementary Table \ref{tab:pval}. For instance, since the enzyme concentration \textit{E} is dependent on the dynamics of the network, the uptake flux $v_1$ can be different between the two models for the same acetate concentration (Equation \ref{eq:flux1}). Consequently, as the enzyme concentration \textit{E} is not part of the closed-form expression for $V_1^{max}$ and $K_1^{ac}(ne)$ in Equation \ref{eq:v1_par}), the difference in the steady state data used for identification can result in a change in the spread (\textcolor{red}{uncertainty}) observed for estimated values of $V_1^{max}$ and $K_1^{ac}(ne)$ (Supplementary Figure). Thus, while quantifiable, the uncertainty due to mismatch in the in vivo and in vitro information will carry over to the estimated parameters.

	However, this issue can be resolved if more in vivo information is used for parameter identification. We first observe this scenario when Equation (\ref{eq:v1_k1cat}), which includes \textit{E}, is used to identify $k_1^{cat}$ and $K_1^{ac}$: these parameters are practically identifiable even when in silico steady state data from a mismatched model is used for identification (Supplementary Figure). We also observe this with the identification of $v_2$ and $v_5$ (Supplementary Figure). For these two fluxes all available and necessary steady state information are part of their identifiability expressions, thereby leaving no room for any uncertainties to propagate from the data through the practical identification process.

	Apart from the mismatch between the in vivo and the in vitro enzyme kinetic rate laws, uncertainty in experimental data also arises due to the presence of noise in the measured experimental data. This noise could be attributed to the measurement error commonly encountered in process analytics. In order to test the robustness of our methodology to practically identify parameters using steady state data with measurement errors, we used in silico experimental data with 5\% additive noise for practical identification, instead of the noise-free data that we have used so far.

	We found that in every case where parameters are structurally identifiable, the noise did not have any effect on the identifiability of the parameters or their estimated values. We also found that inclusion of all necessary data (e.g., the presence or lack thereof of enzyme concentration \textit{E} for $v_1$) can alleviate issues related to using experimental data with errors for identification and estimation: the degrees of identifiability of $V_1^{max}$ and $K_1^{ac}(ne)$ had non-zero standard deviations associated with them, but the degrees of identifiability of $k_1^{cat}$ and $K_1^{ac}$ did not. Using a similar reasoning to the earlier scenario in the presence of mismatches between in vitro and in vivo model, we can say that there is no room for any uncertainties to propagate from the noisy data when all necessary steady state information for identification is available. Thus, both $v_2$ and $v_5$ also did not show any differences in either their degree of identifiability or their estimated parameter values.

	However, for $v_3$, whose parameters are only locally structurally identifiable, we found small non-zero standard deviations in the degrees of identifiabilities (Supplementary Figure) when noisy data is used. Although this was seen due to the differences in the number of data combinations that can estimate positive values for each of the three parameters between different noisy experimental data sets, we observe that the standard deviation in the estimated parameter values for each data, between different samples of noisy experimental data, is small (Supplementary Figure).

	\section{Discussions}\label{sec:discussion}	
	Parameter estimation for kinetic models has always focused on the ability to estimate parameters from existing data without the need for additional experiments, which might not be always possible if parameters are not identifiable from existing experimental data. The presence of noise is typically said to be a significant factor that results in non-identifiability. However, there different reasons for non-identifiability of parameters that we show with our work. First, non-identifiability could be structural to the model used to represent the flux, and cannot be alleviated without reduction in the parameter space. Otherwise, non-identifiability of parameters can be attributed to the lack of information about the dynamics of the system whose parameters are being estimated within the chosen experimental data. The informativeness of experiments can be tied back to their ability to discriminate the dynamics of the system under two or more different input conditions. Thus, the presence of noise only serves to exacerbate the inability of experiments to discriminate the dynamics of the systems.

	Previously, methods have been developed for practical parameter identification and experimental design for kinetic models of metabolism. These methods for experimental design based on practical identification of parameters rely on solving nonlinear least squares problems using optimization approaches that cannot guarantee global optimal solutions \parencite{Raue2009a}, or calculating the Fischer Information Matrix (FIM) to obtain information on the structural and practical identifiability of parameters in kinetic models. Either of these types of methods become computationally cumbersome for models of large genome-scale, or even central carbon scale metabolic networks. Some authors have eschewed deterministic parameter estimation techniques in favour of Bayesian methods based on probabilistic estimation of parameters and experimental design \parencite{Saa2016, Saa2016a} that has the possibility of overcoming some of the issues with the deterministic techniques.

	In this document, we have presented a scalable method to practically identify parameters in kinetic models of metabolism, and use it to design experiments that are minimal and informative for estimating the parameters that does not require solutions to non-convex optimization problems. By establishing identifiability for each flux within a metabolic network individually, we hope to overcome the scalability obstacle. Furthermore, we believe our method offers an algorithmic alternative to determine persistently excitable experiments that can enable identification of all fluxes within a metabolic network. Using a small metabolic network for gluconeogenesis, we have demonstrated that the identifiability of parameters for a given flux is dependent on the position of the flux within the metabolic network. We have also shown the ability to use our analysis to design the minimal number of experiments that are most informative for identifying all fluxes within a metabolic network.

	We find that the identifiability of parameters in kinetic models of metabolism using steady state information is dependent on the kinetic rate law used to model the fluxes within metabolism. The impact of the formulation and nonlinearity of a kinetic rate law expression affecting the practical identifiability of parameters in the expression may not be an unique problem isolated to the system that we are investigating. Complicated expressions for describing fluxes have been extensively used to model observed experimental data for different fluxes in a variety of organisms \parencite{Chassagnole2002a, Peskov2012, VanHeerden2014}. However, authors have favored working with approximate kinetic models of metabolism whose parameters are easily identifiable and estimable instead of trying to establish the identifiability of the parameters used in these models \textcolor{red}{(mention Heijnen papers on resolving identifiability using approximate models here)}.

	We have shown that in some instances (e.g., $v_5$) local practical identifiability could be resolved to obtain global practical identifiability using constraints on the values of the parameters such that they are physically relevant. We have also shown that the structural identifiability of the parameters in any given kinetic rate law model has a bearing on the ability to determine the practical identifiability of parameters using steady state metabolomic, fluxomic and proteomic information. We find that these can sometimes be resolved by reducing the dimension of the parameter space that is being identified: $\theta \in \mathbb{R}^3$ to $\theta \in \mathbb{R}^2$ for both $v_3$ and $v_2$. Additionally, we would also like to point out that discrepancies between in vivo kinetic rate law from which typical experimental data is obtained, and the in vitro rate law used in kinetic models can itself lead to practical parameter non-identifiability or local identifiability. This can lead to uncertainty in parameter estimates made from in vivo experimental data.

	Our work adds to this existing body of work wherein we develop a method for practical identifiability tailored for use with nonlinear enzyme kinetic rate laws that are typically used to model fluxes in metabolic networks. With our work we hope to change the status quo in the application of systems identification techniques for kinetic models of metabolic networks. Our methodology fills the niche gap of experimental design for parameter estimation by providing a way to design informative experiments to obtain data required for parameter estimation by spending the least amount of resources.	
	In the future, we believe our work can be extended and formulated as a mixed integer linear programming problem that can be solved to determine the type and total minimum number of experiments necessary to estimate all parameters in kinetic models of genome-scale metabolic networks.

	\printbibliography	
\end{document}